Malacology: A Programmable Storage System

### 2017-02-24 paper/6_evaluation.tex
paper: wordsmithing and fix colors
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Since this work focuses on the programmability of Malacology, the goal of this
section is to show that the components and subsystems that support the
Malacology interfaces provide reasonable relative performance, as well as to
give examples of the flexibility that Malacology provides to programmers.

@@

We evaluate Malacology by exploring one possible mapping of the ZLog
implementation of CORFU onto Ceph in which we re-use (1) the metadata service
to manage naming and synchronization of the sequencer resource by treating the
resource as an inode, and (2) the monitoring sub-subsystem to distribute and
install application-specific I/O interface required of the CORFU protocol. In
Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the
inode abstraction for implementing the sequencer resource enables load
balancing policies to migrate the sequencer resource in heavy-load situation.

@@

The sequencer is implemented using the File Type interface so that the sequencer state
(a 64-bit integer) is embedded in the inode of a file; this file represents the
sequencer. A total ordering of the log is imposed by the re-use of the
capability service that can be used to grant exclusive access of inode state to
clients. The metadata service is responsible for maintaining exclusivity and
granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
behavior of the system in which a best-effort policy is used.  The two colors
represent points in time that the clients were able to access the resource.
The best-effort policy shows a high degree of interleaving between clients but
the system spends a large portion of time re-distributing the capability,
reducing overall throughput.

@@

Figure~\ref{fig:captp} demonstrates a configurable trade-off between throughput
and latency. In the experiment two clients are run each with a fixed 0.25
second maximum reservation on the capability, and we vary the size of the log
position quota running each configuration for two minutes. The total operations
per second is the combined throughput of the two clients, and the average
latency is number of microseconds required to obtain a new log position. With a
small quota more time is spent exchanging exclusive access, while a large quota
reservation allows clients to experience a much lower latency because they
experience isolated access for a longer period of time.

@@

To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
latency for each client in all experiment configurations. At the 99th
percentile clients accessed the sequencer is less than a millisecond. The CDF
is cropped at the 99.999th percentile due to large outliers that we believe
occur in instances in which the metadata server is performing I/O while it is in the
process of re-distributing the capability to another client.

@@

We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
evaluating the performance of installing and distributing interface updates.
Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
The interfaces are Lua scripts embedded in the cluster map and distributed
using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
time following the Paxos proposal for an interface update until each object storage daemon makes
the update live (the cost of the Paxos proposal is configurable and is
discussed below). The latency measurements were taken on the nodes running Ceph
server daemons, and thus exclude the client round-trip cost. In each of the
experiments 1000 interface updates were observed.

@@

In practice, a storage system implementing CORFU will support a multiplicity of
independent totally-ordered logs for each application.  For this scenario
co-locating sequencers on the same physical node is not ideal but building a
load balancer that can migrate the shared resource ({\it e.g.,}, the resource
that mediates access to the tail of the log) is a time-consuming, non-trivial
task.  It requires building subsystems for migrating resources, monitoring the
workloads, collecting metrics that describe the utilization on the physical
nodes, partitioning resources, maintaining cache coherence, and managing
multiple sequencers. The following experiments demonstrate the feasibility of
using the mechanisms of the Malacology Load Balancing interface to inherit
these features and to alleviate load from overloaded servers.

@@

First, we show how the ZLog service can orchestrate multiple sequencers using
the Malacology Load Balancing interface.
Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
different load balancers as they migrate 3 sequencers (with 4 clients) around
the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
custom load balancer we wrote specifically for sequencers. The CephFS load
balancers are hard-coded into Ceph while Mantle is an API for balancing load
presented in~Mantle and re-implemented on top of
Malacology in this paper.  The increased throughput for the CephFS and Mantle
curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
overloaded servers.

@@

Next, we quantify the performance benefits shown in
Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
perform differently we need to explain the different balancing modes that the
load balancer service uses and how they stress the subsystems that receive and
forward client requests in different ways. In
Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
performance of the balancing mode that CephFS falls into {\it most of the
time}.  CephFS currently has 3 modes for balancing load: CPU mode, workload mode, and
hybrid mode. All three have the same structure for making migration decisions
but vary based on the metric used to calculate load. For this sequencer
workload the 3 different modes all have the same performance, shown in
Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer
falls into the same mode a majority of the time.  The high variation in
performance for the CephFS CPU Mode bar reflects the uncertainty of using
something as dynamic and unpredictable as CPU utilization to make migration
decisions. In addition to the suboptimal performance and unpredictability,
because CephFS balancers all fall into the same mode despite using different
balancers is a problem because it prevents administrators from properly
exploring the balancing state space.

Mantle gives the administrator more control over balancing policies; for the
Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the load
balancing interface to program logic for balancing read-heavy workloads,
resulting in better throughput and stability.  When we did this we also
identified two balancing modes relevant for making migration decisions for
sequencers.

Using Mantle, the administrator can put the load balancer into ``proxy
mode" or ``client mode". In proxy mode one server receives all requests and
farms off the requests to slave servers; the slaves servers do the actual tail
finding operation. In client mode, clients interact directly with the server
that has their sequencer.  These modes are illustrated in
Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
co-located on one physical server -- performance for that mode is shown by the
No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
Mode" in the lower left shows how clients for sequencer 2 are re-directed to
server B. The two ``Proxy mode" diagrams on the right show how clients continue
sending requests to server A even though sequencer 2 has migrated to server B
(the meaning of ``(Half)" is discussed in the next section).

@@

Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
different modes for an environment with only 2 sequencers (again 4 clients
each) and 2 servers. The curves for both sequencers in
Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
Performance of Sequencer 2 decreases because it stayed on the proxy which now
handles all client requests, processes requests for Sequencer 2, and forwards
requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
because distributing the sequencers in this way separates (1) the handling of
the client requests and (2) finding the tail of the log and responding to
clients.  Doing both steps is too heavy weight for one server and sequencers on
slave nodes can go faster if work is split up; this phenomenon is not uncommon
and has been observed in chain replication~\cite{chain_rep}.

Cluster throughput improves at the cost of decreased throughput for Sequencer
2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually
(no balancing phase) and shows that the cluster throughput is worse than the
cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
less throughput than Sequencer 2. In this case, the scatter-gather process used
for cache coherence in the metadata protocols causes strain on the server
housing Sequencer 2 resulting in this uneven performance.

@@

Another factor that affects performance in this environment is how much load is
on each server; these experiments quantify that effect by programming the load
balancing interface to control the amount of load to migrate. We call this
metric a ``migration unit".  Expressing this heuristic is not easily achievable
with outward facing tunable parameters ({\it i.e.} system knobs) but with Mantle's
programmable interface, we can force the load balancer to change its migration
units. To force the balancer into the Proxy Mode (Half) scenario in
Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
load on the current server, we can use: \texttt{ targets[whoami+1] =
mds[whoami]["load"]/2 }.

This code snippet uses globally defined variables and tables from the Mantle
API to send half of the load on the current server (whoami) to the next ranked
server (whoami + 1); the \texttt{targets} array is a globally defined table
that the balancer uses to do the migrations. Alternatively, to migrate all load
a time step, we can remove the division by 2.

@@

Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the
modes using different migration units. Recall that this setup only has 2
sequencers and 2 servers, so performance may be different at scale. Even so, it
is clear that client mode does not perform as well for read-heavy workloads. We
even see a throughput improvement when migrating all load off the first server,
leaving the first server to do administrative tasks (this is common in the
metadata cluster because the first server does a lot of the cache coherence
work) while the second server does all the client work. Proxy mode does the
best in both cases and shows large performance gains when completely decoupling
client request handling and operation processing in Proxy Mode (Full).  The
parameter that controls the migration units helps the administrator control the
sequencer co-location or distribution across the cluster. This trade-off was
explored extensively in the Mantle paper but the experiments we present here
are indicative of an even richer set of states to explore.

@@

Tuning the aggressiveness of the load balancer decision making is also a
trade-off that administrators can control and explore. The balancing phase from
0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
degrees of aggressiveness in making migration decisions; CephFS makes a
decision 10 seconds into the run and throughput jumps to 2500 ops/second within 10 seconds
while Mantle takes more time to stabilize. This conservative behavior is
controlled by programming the balancer to (1) use different conditions for when
to migrate and (2) using a threshold for sustained overload.

>>>>

Since this work focuses on the programmability of Malacology, the goal of this
section is to show that the components and subsystems that support the
Malacology interfaces provide reasonable relative performance, as well as to
give examples of the flexibility that Malacology provides to programmers.  
This section uses a principled approach for evaluating tunables of the
interfaces and the trade-offs we discuss should be acknowledged when building
higher-level services.

@@

We evaluate Malacology by exploring one possible mapping of the ZLog
implementation of CORFU onto Ceph in which we re-use (1) the metadata service
to manage naming and synchronization of the sequencer resource by treating the
resource as an inode, and (2) the monitoring sub-subsystem to distribute and
install application-specific I/O interfaces required of the CORFU protocol. In
Section~\ref{sec:zlog-balancing} we then demonstrate how the re-use of the
inode abstraction for implementing the sequencer resource enables load
balancing policies to migrate the sequencer resource in heavy-load situation.

@@

The sequencer is implemented using the File Type interface so that the sequencer state
(a 64-bit integer) is embedded in the inode of a file.
A total ordering of the log is imposed by the re-use of the
capability service that can be used to grant exclusive access of inode state to
clients. The metadata service is responsible for maintaining exclusivity and
granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
behavior of the system in which a best-effort policy is used.  The two colors
represent points in time that the clients were able to access the resource.
The best-effort policy shows a high degree of interleaving between clients but
the system spends a large portion of time re-distributing the capability,
reducing overall throughput.

@@

Figure~\ref{fig:captp} demonstrates a configurable trade-off between throughput
and latency. In the experiment two clients are run each with a fixed 0.25
second maximum reservation on the capability, and we vary the size of the log
position quota running each configuration for two minutes. The total operations
per second is the combined throughput of the two clients, and the average
latency is the number of microseconds required to obtain a new log position. With a
small quota more time is spent exchanging exclusive access, while a large quota
reservation allows clients to experience a much lower latency because they
experience isolated access for a longer period of time.

@@

To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
latency for each client in all experiment configurations. At the 99th
percentile clients accessed the sequencer in less than a millisecond. The CDF
is cropped at the 99.999th percentile due to large outliers that we believe
occur in instances in which the metadata server is performing I/O while it is in the
process of re-distributing the capability to another client.

@@

We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
evaluating the performance of installing and distributing interface updates.
Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
The interfaces are Lua scripts embedded in the cluster map and distributed
using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
time following the Paxos proposal for an interface update until each object storage daemon makes
the update live (the cost of the Paxos proposal is configurable and is
discussed below). The latency measurements were taken on the nodes running object
server daemons, and thus exclude the client round-trip cost. In each of the
experiments 1000 interface updates were observed.

@@

In practice, a storage system implementing CORFU will support a multiplicity of
independent totally-ordered logs for each application.  For this scenario
co-locating sequencers on the same physical node is not ideal but building a
load balancer that can migrate the shared resource (e.g., the resource
that mediates access to the tail of the log) is a time-consuming, non-trivial
task.  It requires building subsystems for migrating resources, monitoring the
workloads, collecting metrics that describe the utilization on the physical
nodes, partitioning resources, maintaining cache coherence, and managing
multiple sequencers. The following experiments demonstrate the feasibility of
using the mechanisms of the Malacology Load Balancing interface to inherit
these features and to alleviate load from overloaded servers.

@@

First, we show how the ZLog service can orchestrate multiple sequencers using
the Malacology Load Balancing interface.
Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
different load balancers as they migrate 3 sequencers (with 4 clients) around
the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
migrates sequencers using the hard-coded CephFS load balancers, and ``Mantle" uses a
custom load balancer we wrote specifically for sequencers. 
The increased throughput for the CephFS and Mantle
curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
overloaded servers.

@@

Next, we quantify the performance benefits shown in
Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
perform differently we need to explain the different balancing modes that the
load balancer service uses and how they stress the subsystems that receive and
forward client requests in different ways. In
Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
performance of the balancing mode that CephFS falls into {\it most of the
time}.  CephFS currently has 3 modes for balancing load: CPU mode, workload mode, and
hybrid mode. All three have the same structure for making migration decisions
but vary based on the metric used to calculate load. For this sequencer
workload the 3 different modes all have the same performance, shown in
Figure~\ref{fig:mantle-balancer-performance} (a), because the load balancer
falls into the same mode a majority of the time.  The high variation in
performance for the CephFS CPU Mode bar reflects the uncertainty of using
something as dynamic and unpredictable as CPU utilization to make migration
decisions. 
In addition to the suboptimal performance and unpredictability,
another problem is that all the CephFS balancers behave the same. This
prevents administrators from properly
exploring the balancing state space.

Mantle gives the administrator more control over balancing policies; for the
Mantle bar in Figure~\ref{fig:mantle-balancer-performance} (a) we use the Load
Balancing interface to program logic for balancing read-heavy workloads,
resulting in better throughput and stability.  When we did this we also
identified two balancing modes relevant for making migration decisions for
sequencers.

Using Mantle, the administrator can put the load balancer into ``proxy mode" or
``client mode". In proxy mode one server receives all requests and farms off
the requests to slave servers; the slave servers do the actual tail finding
operation. In client mode, clients interact directly with the server that has
their sequencer.  These modes are illustrated in Figure~\ref{fig:mantle-modes}.
``No Balancing" is when all sequencers are co-located on one physical server --
performance for that mode is shown by the ``No Balancing" curve in
Figure~\ref{fig:mantle-balancer-behaviors}. In ``Proxy Mode", clients continue
sending requests to server A even though some of the sequencers have been
migrated to another server. Server A redirects client requests for sequencer 2
to server B.  ``Proxy Mode (Half)" is shown in
Figure~\ref{fig:mantle-balancer-beahviors}; in this scenario, half of the
sequencers have migrated off the first server. Alternatively, ``Proxy Mode
(Full)", which is not pictured, is when all the sequencers migrate off the
first server.  ``Client Mode", shown on the far right of
Figure~\ref{fig:mantle-modes}, shows how clients for sequencer 2 contact server
B without a redirect from server A.

@@

Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
different modes for an environment with only 2 sequencers (again 4 clients
each) and 2 servers. The curves for both sequencers in
Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
Performance of Sequencer 2 decreases because it stayed on the proxy which now
processes requests for Sequencer 2, and forwards
requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
because distributing the sequencers in this way separates (1) the handling of
the client requests and (2) finding the tail of the log and responding to
clients.  Doing both steps is too heavy weight for one server and sequencers on
slave nodes can go faster if work is split up; this phenomenon is not uncommon
and has been observed in chain replication~\cite{chain_rep}.

Cluster throughput improves at the cost of decreased throughput for Sequencer
2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually
(no balancing phase) and shows that the cluster throughput is worse than the
cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
less throughput than Sequencer 1. In this case, the scatter-gather process used
for cache coherence in the metadata protocols causes strain on the server
housing Sequencer 2 resulting in this uneven performance.

@@

Another factor that affects performance in this environment is how much load is
on each server; these experiments quantify that effect by programming the Load
Balancing interface to control the amount of load to migrate. We call this
metric a ``migration unit".  Expressing this heuristic is not easily achievable
with outward facing tunable parameters ({\it i.e.} system knobs) but with Mantle's
programmable interface, we can force the load balancer to change its migration
units. To force the balancer into the Proxy Mode (Half) scenario in
Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
load on the current server, we can use: \texttt{ targets[whoami+1] =
mds[whoami]["load"]/2 }.

This code snippet uses globally defined variables and tables from the Mantle
API~\ref{sevilla:sc15-mantle} to send half of the load on the current server (whoami) to the next ranked
server (whoami + 1); the \texttt{targets} array is a globally defined table
that the balancer uses to do the migrations. Alternatively, to migrate all load
a time step, we can remove the division by 2.

@@

Figure~\ref{fig:mantle-balancer-performance} (b) shows the performance of the
modes using different migration units. Recall that this setup only has 2
sequencers and 2 servers, so performance may be different at scale. Even so, it
is clear that client mode does not perform as well for read-heavy workloads. We
even see a throughput improvement when migrating all load off the first server,
leaving the first server to do administrative tasks (this is common in the
metadata cluster because the first server does a lot of the cache coherence
work) while the second server does all the processing. Proxy mode does the
best in both cases and shows large performance gains when completely decoupling
client request handling and operation processing in Proxy Mode (Full).  The
parameter that controls the migration units helps the administrator control the
sequencer co-location or distribution across the cluster. This trade-off was
explored extensively in the Mantle paper but the experiments we present here
are indicative of an even richer set of states to explore.

@@

Tuning the aggressiveness of the load balancer decision making is also a
trade-off that administrators can control and explore. The balancing phase from
0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
degrees of aggressiveness in making migration decisions; CephFS makes a
decision 10 seconds into the run and throughput jumps to 2500 ops/second 
while Mantle takes more time to stabilize. This conservative behavior is
controlled by programming the balancer to (1) use different conditions for when
to migrate and (2) using a threshold for sustained overload.
